\chapter*{Summary}
The explanation of phenotypes in cancer, such as cell line drug response or
    patient survival, has largely been focussed on genomic alterations (SNPs,
    CNVs, and structural variations). While this approach has generated many
    profound insights into cancer biology, it does not directly make statements
    about the signaling impact those cellular aberrations create. With direct
    measurements much less widely available than gene expression, pathway
    methods (mostly mapping gene expression onto signalling proteins) have so
    far fallen short on delivering actionable evidence. This may in part due to
    lack of robustness, but these approaches are fundamentally at odds with the
    notion of tight post-translational control of signal transduction. A way to
    solve this may be to derive consensus signatures of pathway activity to
    make inferences about signalling, or signatures of specific drugs to
    investigate their interactions.

As a baseline (chapter 2), I investigated how well pathway methods compare to
    driver mutations in terms of explaining cell line drug response. I went on
    to analyse the value of mRNA as a downstream signature of signaling
    activity instead of mapping it to the pathway components by means of the
    previously published SPEED platform. This improved over mapping pathway
    members using Gene Ontology or Reactome, but it considered only sets of
    up-regulated genes defined by multiple arbitrary cutoffs. Hence, I extended
    the data set and created a linear model (chapter 3) that compares
    favourably to gene set as well as state of the art pathway methods in terms
    of recovering driver mutations, cell line drug response, and patient
    survival (chapter 4). To complement this, I investigate how gene expression
    signatures of drugs can be used in conjunction with viability data to
    suggest effective drug combinations where no pathway information is
    available (chapter 5).

To the best of my knowledge, this thesis represents the first comprehensive
    analysis of different pathway methods across primary cohorts and cancer
    cell lines, as well as the first large-scale systematic analysis of drug
    sensitisation that could lead to new drug combinations.
